---
title: "KE5105 - Building Electrical Consumption Forecasting"
output: github_document
---
# Extract, Transform and Load Data 5 - Data Imputation"

# Summary of Findings
* Missing data in large gaps cannot be imputed accurately
  * Value
    * Structural model and spline produces outliers (in value)
  * Variation
    * Structural model, Arima and moving average fail to capture the variation in the time series
* All methods perform similarly for small gaps

# Load libraries
```{r}
library(ggplot2)
library(xts)
library(imputeTS)
```

# Load data
```{r load_data, results='hide'}
sde3_agg_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/test_data/SDE-3.agg.csv", header = TRUE, stringsAsFactors = FALSE)
head(sde3_agg_df)
```

## Convert the Pt_timeStamp strings to POSIX time
```{r convert_data, results='hide'}
sde3_agg_df$Pt_timeStamp <- strptime(sde3_agg_df$Pt_timeStamp, format = "%Y-%m-%d %H:%M:%S", tz="GMT")
head(sde3_agg_df)
```


```{r}
str(sde3_agg_df$Pt_timeStamp[2])
```

## Convert the time series data for plotting
```{r}
ts <- xts(sde3_agg_df$PWM_30min_avg, as.Date(sde3_agg_df$Pt_timeStamp))

#start_time = sde3_agg_df[1, "Pt_timeStamp"]
#end_time = unclass(as.POSIXct(tail(sde3_agg_df, 1)$Pt_timeStamp, origin = as.POSIXct(tz="GMT")))
#ts <- ts(data=sde3_agg_df$PWM_30min_avg, start=c(sde3_agg_df[1, "Pt_timeStamp"], 1), end=c(tail(sde3_agg_df, 1)$Pt_timeStamp, 48), frequency=48)
#ts <- ts(data=sde3_agg_df$PWM_30min_avg, start=c(start_time, 1), end=c(end_time, 48), frequency=48)

head(ts)
```

## Plot the time series data
```{r all_sde3_data, fig.height = 5, fig.width = 10}
autoplot(ts, ylab = "Aggregated PWM", xlab = "Time") + ggtitle("SDE-3 Aggregated PWM")
```

## Get the time series data before the data outage
```{r sde3_data_less, results='hide'}
sde3_less_df = sde3_agg_df[sde3_agg_df$Pt_timeStamp < as.POSIXct("2017-03-31 23:30:00"),]
head(sde3_less_df)
```


```{r}
ts_less <- xts(sde3_less_df$PWM_30min_avg, as.Date(sde3_less_df$Pt_timeStamp))
head(ts)
```

## Plot the time series data
```{r plot_sde3, fig.height = 3, fig.width = 10}
autoplot(ts_less, ylab = "Aggregated PWM", xlab = "Time") +
  ggtitle("SDE-3 Aggregated PWM")
```

## Plot the missing data
```{r plot_sde3_missing, fig.height = 4, fig.width = 10}
plotNA.distribution(sde3_less_df$PWM_30min_avg, cex=.1)
```

## Plot the distribution of the missing data
```{r}
plotNA.distributionBar(sde3_less_df$PWM_30min_avg, breaks = 20)
```

## Plot the distribution of the missing data by gap size
```{r}
plotNA.gapsize(sde3_less_df$PWM_30min_avg)
```

# Impute the missing values using different imputation methods
```{r}
rownames(sde3_less_df) <- sde3_less_df$Pt_timeStamp
head(sde3_less_df["PWM_30min_avg"])
```


```{r}
# Impute the missing values using structural model and Kalman smoothing
imp <- na.kalman(sde3_less_df["PWM_30min_avg"])

# Impute the missing values using ARIMA model and Kalman smoothing
imp_arima <- na.kalman(sde3_less_df["PWM_30min_avg"], model = "auto.arima")

# Impute the missing values using spline interpolation
imp_spline <- na.interpolation(sde3_less_df["PWM_30min_avg"], option = "spline")

# Impute the missing values using moving average
imp_ma <- na.ma(sde3_less_df["PWM_30min_avg"], k=4, weighting="exponential")
```

## Plot the imputed data
```{r plot_struc_imp_all, fig.height = 5, fig.width = 10}
# structural model with kalman smoothing
plotNA.imputations(x.withNA = sde3_less_df$PWM_30min_avg,
                   x.withImputations = imp$PWM_30min_avg,
                   main = "Imputated Values Using Structural Model with Kalman Smoothing",
                   ylim = c(-150, 450), cex = .1)
```


```{r plot_arima_imp_all, fig.height = 5, fig.width = 10}
# arima model with kalman smoothing 
plotNA.imputations(x.withNA = sde3_less_df$PWM_30min_avg,
                   x.withImputations = imp_arima$PWM_30min_avg,
                   main = "Imputed Values Using Arima Model with Kalman Smoothing",
                   ylim = c(-150, 450), cex = .1)
```


```{r plot_spline_imp_all, fig.height = 5, fig.width = 10}
# spline interpolation
plotNA.imputations(x.withNA = sde3_less_df$PWM_30min_avg,
                   x.withImputations = imp_spline$PWM_30min_avg,
                   main = "Imputated Values Using Spline Interpolation",
                   ylim = c(-150, 450), cex = .1)
```


```{r plot_ma_imp_all, fig.height = 5, fig.width = 10}
# exponential moving average
plotNA.imputations(x.withNA = sde3_less_df$PWM_30min_avg,
                   x.withImputations = imp_ma$PWM_30min_avg,
                   main = "Imputed Values Using Exponential Moving Average",
                   ylim = c(-150, 450), cex = .1)
```

Imputation missing data in the larger gaps appear wildly inaccurate for the structural model and spline interpolation. 

## Plot the imputed data for the 1st 5000 observations
```{r plot_struc_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Structural Model With Kalman Smoothing",
                   cex=.1)
```


```{r plot_arima_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp_arima[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Arima Model with Kalman Smoothing", cex = .1)
```

Imputation of missing date in the large gaps fail to capture the variability seen in the time series data.


```{r plot_spline_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp_spline[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Spline Interpolation", cex = .1)
```

Imputation of missing date in the large gaps by spline interpolation captures better the variability seen in the time series data than the other methods.

```{r plot_ma_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp_ma[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Exponential Moving Average", cex = .1)
```

## Plot the imputed data for the 2000-2500 observations
```{r plot_struc_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1950:2550, "PWM_30min_avg"],
                   x.withImputations = imp[1950:2550, "PWM_30min_avg"],
                   main = "Imputed Values Using Structural Model With Kalman Smoothing",
                   ylim = c(30, 140))
```


```{r plot_arima_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1950:2550, "PWM_30min_avg"],
                   x.withImputations = imp_arima[1950:2550, "PWM_30min_avg"],
                   main = "Imputed Values Using Arima Model With Kalman Smoothing",
                   ylim = c(30, 140))
```


```{r plot_spline_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1950:2550, "PWM_30min_avg"],
                   x.withImputations = imp_spline[1950:2550, "PWM_30min_avg"],
                   main = "Imputed Values Using Spline Interpolation",
                   ylim = c(30, 140))
```


```{r plot_ma_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[1950:2550, "PWM_30min_avg"],
                   x.withImputations = imp_ma[1950:2550, "PWM_30min_avg"],
                   main = "Imputed Values Using Exponential Moving Average",
                   ylim = c(30, 140))
```

Imputation of missing data in the smaller gaps appear to be quite accurate.

## Plot the imputed data for the 3500-4000 observations
```{r plot_struc_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[3500:4000, "PWM_30min_avg"], x.withImputations = imp[3500:4000, "PWM_30min_avg"])
```


```{r plot_arima_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[3500:4000, "PWM_30min_avg"], x.withImputations = imp_arima[3500:4000, "PWM_30min_avg"])
```


```{r plot_spline_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[3500:4000, "PWM_30min_avg"], x.withImputations = imp_spline[3500:4000, "PWM_30min_avg"])
```


```{r plot_ma_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[3500:4000, "PWM_30min_avg"], x.withImputations = imp_ma[3500:4000, "PWM_30min_avg"])
```

## Plot the imputed data for the 4200-5000 observations
```{r plot_struc_imp4, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[4230:5000, "PWM_30min_avg"], x.withImputations = imp[4230:5000, "PWM_30min_avg"])
```


```{r plot_arima_imp4, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[4230:5000, "PWM_30min_avg"], x.withImputations = imp_arima[4230:5000, "PWM_30min_avg"])
```


```{r plot_spline_imp4, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[4230:5000, "PWM_30min_avg"], x.withImputations = imp_spline[4230:5000, "PWM_30min_avg"])
```


```{r plot_ma_imp4, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[4230:5000, "PWM_30min_avg"], x.withImputations = imp_ma[4230:5000, "PWM_30min_avg"])
```

## Plot the imputed data for the 5000-6000 observations
```{r plot_struc_imp5, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[5000:6000, "PWM_30min_avg"], x.withImputations = imp[5000:6000, "PWM_30min_avg"])
```


```{r plot_arima_imp5, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[5000:6000, "PWM_30min_avg"], x.withImputations = imp_arima[5000:6000, "PWM_30min_avg"])
```


```{r plot_spline_imp5, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[5000:6000, "PWM_30min_avg"], x.withImputations = imp_spline[5000:6000, "PWM_30min_avg"])
```


```{r plot_ma_imp5, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[5000:6000, "PWM_30min_avg"], x.withImputations = imp_ma[5000:6000, "PWM_30min_avg"])
```

## Plot the imputed data at 14000 observations
```{r plot_struc_imp6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[13500:14500, "PWM_30min_avg"], x.withImputations = imp[13500:14500, "PWM_30min_avg"])
```


```{r plot_arima_imp6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[13500:14500, "PWM_30min_avg"], x.withImputations = imp_arima[13500:14500, "PWM_30min_avg"])
```


```{r plot_spline_imp6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[13500:14500, "PWM_30min_avg"], x.withImputations = imp_spline[13500:14500, "PWM_30min_avg"])
```


```{r plot_ma_imp6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[13500:14500, "PWM_30min_avg"], x.withImputations = imp_ma[13500:14500, "PWM_30min_avg"])
```

## Plot the imputed data for the >30000 observations
```{r plot_struc_imp7, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[31900:32200, "PWM_30min_avg"],
                   x.withImputations = imp[31900:32200, "PWM_30min_avg"],
                   main = "Imputed Values Using Structural Model with Kalman Smoothing",
                   ylim = c(-25, 130))
```


```{r plot_arima_imp7, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[31900:32200, "PWM_30min_avg"],
                   x.withImputations = imp_arima[31900:32200, "PWM_30min_avg"],
                   main = "Imputed Values Using Arima Model with Kalman Smoothing",
                   ylim = c(-25, 130))
```


```{r plot_spline_imp7, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[31900:32200, "PWM_30min_avg"],
                   x.withImputations = imp_spline[31900:32200, "PWM_30min_avg"],
                   main = "Imputed Values Using Spline Interpolation",
                   ylim = c(-25, 130))
```


```{r plot_ma_imp7, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = sde3_less_df[31900:32200, "PWM_30min_avg"],
                   x.withImputations = imp_ma[31900:32200, "PWM_30min_avg"],
                   main = "Imputed Values Using Exponential Moving Average",
                   ylim = c(-25, 130))
```

