---
title: "KE5105 - Building Electrical Consumption Forecasting"
output: github_document
---
# Extract, Transform and Load Data 5 - Data Imputation MD1 & MD6

# Summary of Findings
* Missing data occurs periodically at the end of each month.
* All methods perform similarly for small gaps.
* Test for missing completely at random (MCAR) patterns show some evidence that the missing data is not MCAR in MD1.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

# Load libraries
```{r}
library(ggplot2)
library(xts)
library(imputeTS)
library(MissMech)
```

# MD1
# Load data
```{r load_data, results='hide'}
md1_agg_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/processed_bldg_data/MD1.csv", header = TRUE, stringsAsFactors = FALSE)
head(md1_agg_df)
```

## Convert the Pt_timeStamp strings to POSIX time
```{r convert_data, results='hide'}
md1_agg_df$Pt_timeStamp <- strptime(md1_agg_df$Pt_timeStamp, format = "%Y-%m-%d %H:%M:%S", tz="GMT")
head(md1_agg_df)
```


```{r}
str(md1_agg_df$Pt_timeStamp[2])
```

## Convert the time series data for plotting
```{r}
ts <- xts(md1_agg_df$PWM_30min_avg, as.Date(md1_agg_df$Pt_timeStamp))
head(ts)
```

## Plot the time series data
```{r all_md1_data, fig.height = 5, fig.width = 10}
autoplot(ts, ylab = "Aggregated PWM", xlab = "Time") + ggtitle("SDE-3 Aggregated PWM")
```

## Get the time series data after the data outage
```{r md1_data_less, results='hide'}
md1_less_df = md1_agg_df[md1_agg_df$Pt_timeStamp > as.POSIXct("2018-01-01 00:00:00", tz = "GMT") &
                           md1_agg_df$Pt_timeStamp < as.POSIXct("2018-12-02 23:30:00", tz = "GMT"),]
head(md1_less_df)
```


```{r}
ts_less <- xts(md1_less_df$PWM_30min_avg, as.Date(md1_less_df$Pt_timeStamp))
head(ts_less)
```

## Plot the time series data
```{r plot_md1, fig.height = 3, fig.width = 10}
autoplot(ts_less, ylab = "Aggregated PWM", xlab = "Time") +
  ggtitle("MD1 Aggregated PWM from 1 Jan 2018")
```

## Plot the missing data
```{r plot_md1_missing, fig.height = 4, fig.width = 10}
plotNA.distribution(md1_less_df$PWM_30min_avg, cex=.1)
```

Many of the NAs appear at the end of each month (below) for the time period 2300H to 0000H (3 observations).

```{r}
md1_less_df[is.na(md1_less_df$PWM_30min_avg) & (md1_less_df$Pt_timeStamp > as.POSIXct("2018-03-25 00:00:00")),
            "Pt_timeStamp"]
```

## Plot the distribution of the missing data
```{r}
plotNA.distributionBar(md1_less_df$PWM_30min_avg, breaks = 11)
```

## Plot the distribution of the missing data by gap size
```{r, fig.height = 4, fig.width = 10}
plotNA.gapsize(md1_less_df$PWM_30min_avg, byTotalNA = TRUE)
```

## Missing data statistics
```{r}
statsNA(md1_less_df$PWM_30min_avg, bins = 11)
```

# Impute the missing values using different imputation methods
```{r}
# Impute the missing values using structural model and Kalman smoothing
imp <- na.kalman(md1_less_df["PWM_30min_avg"])

# Impute the missing values using ARIMA model and Kalman smoothing
imp_arima <- na.kalman(md1_less_df["PWM_30min_avg"], model = "auto.arima")

# Impute the missing values using spline interpolation
imp_spline <- na.interpolation(md1_less_df["PWM_30min_avg"], option = "spline")

# Impute the missing values using moving average
imp_ma <- na.ma(md1_less_df["PWM_30min_avg"], k=4, weighting="exponential")
```

## Plot the imputed data
```{r plot_struc_imp_all, fig.height = 5, fig.width = 10}
# structural model with kalman smoothing
plotNA.imputations(x.withNA = md1_less_df$PWM_30min_avg,
                   x.withImputations = imp$PWM_30min_avg,
                   main = "Imputated Values Using Structural Model with Kalman Smoothing",
                   ylim = c(-200, 550), cex = .1)
```


```{r plot_arima_imp_all, fig.height = 5, fig.width = 10}
# arima model with kalman smoothing 
plotNA.imputations(x.withNA = md1_less_df$PWM_30min_avg,
                   x.withImputations = imp_arima$PWM_30min_avg,
                   main = "Imputed Values Using Arima Model with Kalman Smoothing",
                   ylim = c(-200, 550), cex = .1)
```


```{r plot_spline_imp_all, fig.height = 5, fig.width = 10}
# spline interpolation
plotNA.imputations(x.withNA = md1_less_df$PWM_30min_avg,
                   x.withImputations = imp_spline$PWM_30min_avg,
                   main = "Imputated Values Using Spline Interpolation",
                   ylim = c(-200, 550), cex = .1)
```


```{r plot_ma_imp_all, fig.height = 5, fig.width = 10}
# exponential moving average
plotNA.imputations(x.withNA = md1_less_df$PWM_30min_avg,
                   x.withImputations = imp_ma$PWM_30min_avg,
                   main = "Imputed Values Using Exponential Moving Average",
                   ylim = c(-200, 550), cex = .1)
```

Imputation missing data in the larger gaps appear wildly inaccurate for the arima model. 

## Plot the imputed data for the 1st 5000 observations
```{r plot_struc_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Structural Model With Kalman Smoothing",
                   ylim = c(-200, 550), cex=.1)
```


```{r plot_arima_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp_arima[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Arima Model with Kalman Smoothing", ylim = c(-200, 550), cex = .1)
```

Imputation of missing date in the large gaps fail to capture the variability seen in the time series data.


```{r plot_spline_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp_spline[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Spline Interpolation", ylim = c(-200, 550), cex = .1)
```

Imputation of missing date in the large gaps by spline interpolation captures better the variability seen in the time series data than the other methods.

```{r plot_ma_imp1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[1500:5000, "PWM_30min_avg"],
                   x.withImputations = imp_ma[1500:5000, "PWM_30min_avg"],
                   main = "Imputed Values Using Exponential Moving Average", ylim = c(-200, 550), cex = .1)
```

## Plot the imputed data for a large gap
```{r plot_struc_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[2550:2800, "PWM_30min_avg"],
                   x.withImputations = imp[2550:2800, "PWM_30min_avg"],
                   main = "Imputed Values Using Structural Model With Kalman Smoothing",
                   ylim = c(-200, 550), cex = .1)
```

In addition, imputation of data in the large gap fails to capture the variation in the data.

```{r plot_arima_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[2550:2800, "PWM_30min_avg"],
                   x.withImputations = imp_arima[2550:2800, "PWM_30min_avg"],
                   main = "Imputed Values Using Arima Model With Kalman Smoothing",
                   ylim = c(-200, 550), cex = .1)
```


```{r plot_spline_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[2550:2800, "PWM_30min_avg"],
                   x.withImputations = imp_spline[2550:2800, "PWM_30min_avg"],
                   main = "Imputed Values Using Spline Interpolation",
                   ylim = c(-200, 550), cex = .1)
```


```{r plot_ma_imp2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[2550:2800, "PWM_30min_avg"],
                   x.withImputations = imp_ma[2550:2800, "PWM_30min_avg"],
                   main = "Imputed Values Using Exponential Moving Average",
                   ylim = c(-200, 550), cex = .1)
```

Imputation of missing data in the smaller gaps appear to be quite accurate (below).

## Plot the imputed data at around the 4000th observation
```{r plot_struc_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[4300:4400, "PWM_30min_avg"], x.withImputations = imp[4300:4400, "PWM_30min_avg"])
```


```{r plot_arima_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[4300:4400, "PWM_30min_avg"], x.withImputations = imp_arima[4300:4400, "PWM_30min_avg"])
```


```{r plot_spline_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[4300:4400, "PWM_30min_avg"], x.withImputations = imp_spline[4300:4400, "PWM_30min_avg"])
```


```{r plot_ma_imp3, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md1_less_df[4300:4400, "PWM_30min_avg"], x.withImputations = imp_ma[4300:4400, "PWM_30min_avg"])
```

## Prepare the data for MCAR test.
```{r}
md1_less_df$time <- as.numeric(rownames(md1_less_df))
head(md1_less_df)
```

## Check the normality of the distribution.
```{r plot_distribution_density}
plot(density(md1_less_df$PWM_30min_avg, na.rm = TRUE))
```

The distribution does not look normal.

## Perform the Shapiro-Wilks normality test. Limited to 5000 samples in test data due to test constraints.
```{r}
shapiro.test(md1_less_df[1:5000, c("PWM_30min_avg")])
```

The distribution of the data is significantly different from normal distribution.

## Perform the test for MCAR (missing completely at random) non-normal data distribution

```{r}
out <- TestMCARNormality(data=md1_less_df[, c("time", "PWM_30min_avg")], method = "Nonparametric")
```


```{r}
summary(out)
```

There is evidence to reject the null hypothesis (the missing data pattern is **not** MCAR).

# MD6
# Load data
```{r load_data1, results='hide'}
md6_agg_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/processed_bldg_data/MD6.csv", header = TRUE, stringsAsFactors = FALSE)
head(md6_agg_df)
```

## Convert the Pt_timeStamp strings to POSIX time
```{r convert_data1, results='hide'}
md6_agg_df$Pt_timeStamp <- strptime(md6_agg_df$Pt_timeStamp, format = "%Y-%m-%d %H:%M:%S", tz="GMT")
head(md6_agg_df)
```


```{r}
str(md6_agg_df$Pt_timeStamp[2])
```

## Convert the time series data for plotting
```{r}
ts <- xts(md6_agg_df$PWM_30min_avg, as.Date(md6_agg_df$Pt_timeStamp))
head(ts)
```

## Plot the time series data
```{r all_md6_data, fig.height = 5, fig.width = 10}
autoplot(ts, ylab = "Aggregated PWM", xlab = "Time") + ggtitle("MD6 Aggregated PWM")
```

## Get the time series data after the data outage
```{r md6_data_less, results='hide'}
md6_less_df = md6_agg_df[md6_agg_df$Pt_timeStamp > as.POSIXct("2018-01-01 00:00:00", tz = "GMT") &
                           md6_agg_df$Pt_timeStamp < as.POSIXct("2018-12-02 23:30:00", tz = "GMT"),]
head(md6_less_df)
```


```{r}
ts_less <- xts(md6_less_df$PWM_30min_avg, as.Date(md6_less_df$Pt_timeStamp))
head(ts_less)
```

## Plot the time series data
```{r plot_md6, fig.height = 3, fig.width = 10}
autoplot(ts_less, ylab = "Aggregated PWM", xlab = "Time") +
  ggtitle("MD6 Aggregated PWM from 1 Jan 2018")
```

## Plot the missing data
```{r plot_md6_missing, fig.height = 4, fig.width = 10}
plotNA.distribution(md6_less_df$PWM_30min_avg, cex=.1)
```

Many of the NAs appear at the end of each month (below) for the time period 2330H to 0000H (2 observations).

```{r}
md6_less_df[is.na(md6_less_df$PWM_30min_avg), "Pt_timeStamp"]
```

## Plot the distribution of the missing data
```{r}
plotNA.distributionBar(md6_less_df$PWM_30min_avg, breaks = 11)
```

## Plot the distribution of the missing data by gap size
```{r, fig.height = 4, fig.width = 10}
plotNA.gapsize(md6_less_df$PWM_30min_avg, byTotalNA = TRUE)
```

## Missing data statistics
```{r}
statsNA(md6_less_df$PWM_30min_avg, bins = 11)
```

# Impute the missing values using different imputation methods
```{r}
# Impute the missing values using structural model and Kalman smoothing
imp <- na.kalman(md6_less_df["PWM_30min_avg"])

# Impute the missing values using ARIMA model and Kalman smoothing
imp_arima <- na.kalman(md6_less_df["PWM_30min_avg"], model = "auto.arima")

# Impute the missing values using spline interpolation
imp_spline <- na.interpolation(md6_less_df["PWM_30min_avg"], option = "spline")

# Impute the missing values using moving average
imp_ma <- na.ma(md6_less_df["PWM_30min_avg"], k=4, weighting="exponential")
```

## Plot the imputed data
```{r plot_struc_imp1md6, fig.height = 5, fig.width = 10}
# structural model with kalman smoothing
plotNA.imputations(x.withNA = md6_less_df[1450:1550, "PWM_30min_avg"],
                   x.withImputations = imp[1450:1550, "PWM_30min_avg"],
                   main = "Imputated Values Using Structural Model with Kalman Smoothing",
                   ylim = c(500, 1000), cex = .1)
```


```{r plot_arima_imp1md6, fig.height = 5, fig.width = 10}
# arima model with kalman smoothing 
plotNA.imputations(x.withNA = md6_less_df[1450:1550, "PWM_30min_avg"],
                   x.withImputations = imp_arima[1450:1550, "PWM_30min_avg"],
                   main = "Imputed Values Using Arima Model with Kalman Smoothing",
                   ylim = c(500, 1000), cex = .1)
```


```{r plot_spline_imp1md6, fig.height = 5, fig.width = 10}
# spline interpolation
plotNA.imputations(x.withNA = md6_less_df[1450:1550, "PWM_30min_avg"],
                   x.withImputations = imp_spline[1450:1550, "PWM_30min_avg"],
                   main = "Imputated Values Using Spline Interpolation",
                   ylim = c(500, 1000), cex = .1)
```


```{r plot_ma_imp1md6, fig.height = 5, fig.width = 10}
# exponential moving average
plotNA.imputations(x.withNA = md6_less_df[1450:1550, "PWM_30min_avg"],
                   x.withImputations = imp_ma[1450:1550, "PWM_30min_avg"],
                   main = "Imputed Values Using Exponential Moving Average",
                   ylim = c(500, 1000), cex = .1)
```

## Plot the imputed data for another small gap
```{r plot_struc_imp2md6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md6_less_df[10500:10550, "PWM_30min_avg"],
                   x.withImputations = imp[10500:10550, "PWM_30min_avg"],
                   main = "Imputed Values Using Structural Model With Kalman Smoothing",
                   ylim = c(500, 1000), cex=.1)
```


```{r plot_arima_imp2md6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md6_less_df[10500:10550, "PWM_30min_avg"],
                   x.withImputations = imp_arima[10500:10550, "PWM_30min_avg"],
                   main = "Imputed Values Using Arima Model with Kalman Smoothing", ylim = c(500, 1000), cex = .1)
```


```{r plot_spline_imp2md6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md6_less_df[10500:10550, "PWM_30min_avg"],
                   x.withImputations = imp_spline[10500:10550, "PWM_30min_avg"],
                   main = "Imputed Values Using Spline Interpolation", ylim = c(500, 1000), cex = .1)
```


```{r plot_ma_imp2md6, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = md6_less_df[10500:10550, "PWM_30min_avg"],
                   x.withImputations = imp_ma[10500:10550, "PWM_30min_avg"],
                   main = "Imputed Values Using Exponential Moving Average", ylim = c(500, 1000), cex = .1)
```

## Prepare the data for MCAR test.
```{r}
md6_less_df$time <- as.numeric(rownames(md6_less_df))
head(md6_less_df)
```

## Check the normality of the distribution.
```{r plot_distribution_density_md6}
plot(density(md6_less_df$PWM_30min_avg, na.rm = TRUE))
```

The distribution does not look normal.

## Perform the Shapiro-Wilks normality test. Limited to 5000 samples in test data.
```{r}
shapiro.test(md6_less_df[1:5000, c("PWM_30min_avg")])
```

The distribution of the data is significantly different from normal distribution.

## Perform the test for MCAR (missing completely at random) assuming non-normal data distribution
```{r}
out <- TestMCARNormality(data=md6_less_df[, c("time", "PWM_30min_avg")], method = "Nonparametric")
```


```{r}
summary(out)
```

There is insufficient evidence to reject the null hypothesis (the missing data pattern **is MCAR**).
