---
title: "KE5105 - Building Electrical Consumption Forecasting"
output: github_document
---

# Extract, Transform and Load Data 7 - Data Imputation

# Summary of Findings
* Data imputation was performed on the 10 test periods of SDE-3 data with simulated missing completely at random (MCAR) data using :
    * structural model with Kalman smoothing
    * arima model with Kalman smoothing
    * spline interpolation
    * exponential moving average (k=4)
* Outliers in the rmse were found in arima for lambda = 0.15
* 1-way anova test found
    * for lambda = 0.1, no significant differences
    * lambda = 0.15 show significant differences (arima excluded)
    * lambda = 0.2, 0.25 show significant differences
* Multiple pairwise t-tests found the following relationships between rmse:
    * for lambda = 0.1, no significant differences
    * for lambda = 0.2,
        * spline, ma > arima
        * spline, ma > struc
    * for lambda = 0.25
        * spline > ma, arima, struc
        * ma > arima, struct
    * for lambda = 0.15, 
        * spline > struc
  
# Load libraries
```{r}
library(ggplot2)
library(xts)
library(imputeTS)
```

# Load data
```{r load_data, results='hide'}
# SDE-3 time series data
sde3_agg_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/test_data/SDE-3.agg.csv", header = TRUE, stringsAsFactors = FALSE)

# A list of 10 time series datasets with simulated missing data
missing_list <- readRDS("sim_missing_data.rds")
```


```{r}
length(missing_list)
```


```{r}
num_lambdas = length(missing_list)
num_datasets = length(missing_list[[1]])-1  # each list has lambda, missing1, ..., missing<n>

results = vector('list', num_lambdas * num_datasets) 
idx = 1
for (i in missing_list) {

  for (j in 1:num_datasets) {
  # test data
  data_with_na = i[[j+1]]$data
  data_true = sde3_agg_df[i[[j+1]]$row:(i[[j+1]]$row+i[[j+1]]$size-1), ]$PWM_30min_avg
  na_true = data_true[i[[j+1]]$na.ind]
  
  # Impute the missing values using structural model and Kalman smoothing
  imp_struct <- na.kalman(data_with_na)
  # Impute the missing values using ARIMA model and Kalman smoothing
  imp_arima <- na.kalman(data_with_na, model = "auto.arima")
  # Impute the missing values using spline interpolation
  imp_spline <- na.interpolation(data_with_na, option = "spline")
  # Impute the missing values using moving average
  imp_ma <- na.ma(data_with_na, k=4, weighting="exponential")  

  # Extract all the imputed na values
  na_imp_struct = imp_struct[i[[j+1]]$na.ind]
  na_imp_arima = imp_arima[i[[j+1]]$na.ind]
  na_imp_spline = imp_spline[i[[j+1]]$na.ind]
  na_imp_ma = imp_ma[i[[j+1]]$na.ind]
  
  # compute rmse
  rmse_struct = sqrt(mean((na_imp_struct - na_true)^2))
  rmse_arima = sqrt(mean((na_imp_arima - na_true)^2))
  rmse_spline = sqrt(mean((na_imp_spline - na_true)^2))
  rmse_ma = sqrt(mean((na_imp_ma - na_true)^2))
  
  #Update results
  results[[idx]] <- c(i$lambda, rmse_struct, rmse_arima, rmse_spline, rmse_ma)
  idx = idx + 1
  }
}  
```


```{r}
results_df = as.data.frame(matrix(unlist(results), nrow = length(results), byrow = TRUE))
colnames(results_df) = c("lambda", "struc", "arima", "spline", "ma")
results_df
```


```{r}
colMeans(results_df)
# same as apply(results_df, 2, mean)
```

```{r}
apply(results_df, 2, sd)
```


```{r}
aggregate(results_df[, 2:5], list(lambda = results_df$lambda), mean)
```


```{r}
aggregate(results_df[, 2:5], list(lambda = results_df$lambda), sd)
```
There are **outliers** in the RMSE for ARIMA imputation for test samples #6 and #10.

## Check the outliers in the ARIMA imputation.
```{r}
num_lambdas = 1
num_datasets = 2

results = vector('list', num_lambdas * num_datasets) 
idx = 1
  #for (j in 1:num_datasets) {
  for (j in list(7, 11, 5)) {
  # test data
  data_with_na = missing_list[[2]][[j]]$data
  data_true = sde3_agg_df[missing_list[[2]][[j]]$row:(missing_list[[2]][[j]]$row+missing_list[[2]][[j]]$size-1), ]$PWM_30min_avg
  na_true = data_true[missing_list[[2]][[j]]$na.ind]
  
  # Impute the missing values using ARIMA model and Kalman smoothing
  imp_arima <- na.kalman(data_with_na, model = "auto.arima")

  # Extract all the imputed na values
  na_imp_arima = imp_arima[missing_list[[2]][[j]]$na.ind]

  # compute rmse
  rmse_arima = sqrt(mean((na_imp_arima - na_true)^2))

  #Update results
  results[[idx]] <- c(lambda=missing_list[[2]]$lambda, rmse=rmse_arima, imp=list(imp_arima), data.withNA=list(data_with_na))
  idx = idx + 1
}  
```


```{r plot_arima_imp_outliers1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = results[[1]]$data.withNA,
                   x.withImputations = results[[1]]$imp,
                   main = "Imputated Values with High RMSE Using ARIMA Model with Kalman Smoothing",
                   ylim = c(-10, 150), cex = .1)
```

```{r plot_arima_imp_outliers2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = results[[2]]$data.withNA,
                   x.withImputations = results[[2]]$imp,
                   main = "Poorly Imputated Values with High RMSE Using ARIMA Model with Kalman Smoothing",
                   ylim = c(-10, 100), cex = .1)
```


```{r plot_arima_imp_normal, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = results[[3]]$data.withNA,
                   x.withImputations = results[[3]]$imp,
                   main = "Imputated Values with Low RMSE Using ARIMA Model with Kalman Smoothing",
                   ylim = c(30, 150), cex = .1)
```

## Check for significance in the difference in the means of the RMSE.

### One-way ANOVA

#### Determine whether the means from the 4 imputation methods are equal.
```{r}
results_long_df <- reshape(results_df, varying = c('struc', 'arima', 'spline', 'ma'),
                           timevar = "method", v.names='RMSE',
                           times=c('struc', 'arima', 'spline', 'ma'), direction="long")
results_long_df$method <- as.factor(results_long_df$method)
head(results_long_df)
```


```{r boxplots}
for (i in list(.1, .15, .2, .25)) {
  plot(RMSE ~ method, data = results_long_df[results_long_df$lambda == i, ], main = paste("lambda =", toString(i)))
}
```


```{r}
for (i in list(.1, .2, .25)) {
  cat("lambda = ", i, "\n")
  res = aov(RMSE ~ method, data = results_long_df[results_long_df$lambda == i, ])
  print(summary(res))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```

For lambda = .1, the anova test found **insufficient evidence** to show that the differences are significant.
for lambda = .2, .25, the anova tests show **significant** differences.

For lambda = .15, exclude arima from the anova tests because it has outliers in the rmse.
```{r}
res = aov(RMSE ~ method, data = results_long_df[(results_long_df$lambda == .15) & (results_long_df$method != 'arima') , ])
print(summary(res))
```

The anova test for lambda = .15 show **significant** differences between struc, spline and ma.

### Pairwise t-test for lambda .1, .2, .25 (exclude .15 due to outliers in ARIMA)

Pairwise multiple comparisons tests involve the computation of a p-value for each pair of the compared group. The p-value represents the risk of stating that an effect is statistically significant while this is not true. As the number of pairwise comparisons increases, and therefore the number of p-values, it becomes more likely to detect significant effects which are due to chance in reality.

To deal with this problem, multiple pairwise comparisons tests involve p-value corrections: p-values are penalized (= their value is increased) as the number of comparisons increase.

#### Results Summary
for lambda = 0.1, no significant differences.
for lambda = 0.2, spline > arima, ma > arima, ma > struc, spline > struc
for lambda = 0.25, spline > ma, spline > arima, ma > arima, ma > struc, spline > struc

for lambda = .15, spline > struc (arima excluded from test)

```{r}
for (i in list(.1, .2, .25)) {
  cat("lambda = ", i, "\n")
  print(pairwise.t.test(results_long_df[results_long_df$lambda == i,]$RMSE,
                        results_long_df[results_long_df$lambda == i,]$method,
                        p.adjust = "bonferroni"))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```

for lambda = .15, exclude arima from test
```{r}
print(pairwise.t.test(results_long_df[(results_long_df$lambda == .15) & (results_long_df$method != "arima"),]$RMSE,
                      results_long_df[(results_long_df$lambda == .15) & (results_long_df$method != "arima"),]$method,
                      p.adjust = "bonferroni"))
```

