---
title: "KE5105 - Building Electrical Consumption Forecasting"
output: github_document
---

# Extract, Transform and Load Data 7 - Data Imputation - MD1 and MD6

# Summary of Findings
* Data imputation was performed on the 20 test periods of MD1 and MD6 data with simulated missing completely at random (MCAR) data using :
    * structural model with Kalman smoothing
    * arima model with Kalman smoothing
    * spline interpolation
    * exponential moving average (k=4)
* Outliers in the rmse were found in arima for all lambda values
* 1-way anova test found
    * for combined md1, md6 data, no significant differences
    * for md1 data, no significant differences
    * for md6 data, significant differences for all lambda
* Multiple pairwise t-tests found the following relationships between rmse:
    * for combined md1, md6 data, no significant differences
    * for md1 data, no significant differences
    * for md6 data :
      * for lambda=.1, ma > struc
      * for lambda=.15, ma > struc, spline
      * for lambda=.2, ma, spline > struc
      * for lambda=.25, ma > struc, spline

# Load libraries
```{r}
library(ggplot2)
library(xts)
library(imputeTS)
```

# Load data
```{r load_data, results='hide'}
# Time series data
md1_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/processed_bldg_data/MD1.csv", header = TRUE, stringsAsFactors = FALSE)
md6_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/processed_bldg_data/MD6.csv", header = TRUE, stringsAsFactors = FALSE)

# A list of 10 time series datasets with simulated missing data
missing_list_md1 <- readRDS("/home/tkokkeng/Documents/KE5105/ETL/source/md1-data/sim_missing_data.md1.rds")
missing_list_md6 <- readRDS("/home/tkokkeng/Documents/KE5105/ETL/source/md6-data/sim_missing_data.md6.rds")
```


```{r}
cat("length MD1 data = ", length(missing_list_md1), "\n", "length MD6 data = ", length(missing_list_md6))
```


```{r}
computeResults <- function(bldg_name, nresults, missing_list, data_df) {
  
  #results = vector('list', num_lambdas * num_datasets)
  num_datasets = length(missing_list[[1]])-1
  results = vector('list', nresults)
  idx = 1
  for (i in missing_list) {

    for (j in 1:num_datasets) {
      # test data
      data_with_na = i[[j+1]]$data
      data_true = data_df[i[[j+1]]$row:(i[[j+1]]$row+i[[j+1]]$size-1), ]$PWM_30min_avg
      na_true = data_true[i[[j+1]]$na.ind]
      
      # Impute the missing values using structural model and Kalman smoothing
      imp_struct <- na.kalman(data_with_na)
      # Impute the missing values using ARIMA model and Kalman smoothing
      imp_arima <- na.kalman(data_with_na, model = "auto.arima")
      # Impute the missing values using spline interpolation
      imp_spline <- na.interpolation(data_with_na, option = "spline")
      # Impute the missing values using moving average
      imp_ma <- na.ma(data_with_na, k=4, weighting="exponential")  
    
      # Save the imputed data to file (.csv)
      for (k in list("imp_struct", "imp_arima", "imp_spline", "imp_ma")) {
        df = data.frame(eval(parse(text = k)))
        colnames(df) = c('PWM_30min_avg')
        write.csv(df,
                  file = paste(bldg_name, "_", k, "_", i$lambda, "_", j, ".csv", sep = ""),
                  row.names = FALSE)
      }
      
      # Extract all the imputed na values
      na_imp_struct = imp_struct[i[[j+1]]$na.ind]
      na_imp_arima = imp_arima[i[[j+1]]$na.ind]
      na_imp_spline = imp_spline[i[[j+1]]$na.ind]
      na_imp_ma = imp_ma[i[[j+1]]$na.ind]
      
      # compute rmse
      rmse_struct = sqrt(mean((na_imp_struct - na_true)^2))
      rmse_arima = sqrt(mean((na_imp_arima - na_true)^2))
      rmse_spline = sqrt(mean((na_imp_spline - na_true)^2))
      rmse_ma = sqrt(mean((na_imp_ma - na_true)^2))
      
      #Update results
      results[[idx]] <- c(i$lambda, rmse_struct, rmse_arima, rmse_spline, rmse_ma)
      idx = idx + 1
    }
  }
  return(results)
}
```


```{r}
# Each list has lambda, missing1, ..., missing<n>
nresults_md1 <- length(missing_list_md1) * (length(missing_list_md1[[1]])-1)
nresults_md6 <- length(missing_list_md6) * (length(missing_list_md6[[1]])-1)
results_md1 <- computeResults("md1", nresults_md1, missing_list_md1, md1_df)
results_md6 <- computeResults("md6", nresults_md6, missing_list_md6, md6_df)
results <- c(results_md1, results_md6)
```


```{r}
results_df = as.data.frame(matrix(unlist(results), nrow = length(results), byrow = TRUE))
colnames(results_df) = c("lambda", "struc", "arima", "spline", "ma")
results_df
```


```{r}
colMeans(results_df)
# same as apply(results_df, 2, mean)
```

```{r}
apply(results_df, 2, sd)
```


```{r}
aggregate(results_df[, 2:5], list(lambda = results_df$lambda), mean)
```


```{r}
aggregate(results_df[, 2:5], list(lambda = results_df$lambda), sd)
```
There are **outliers** in the RMSE for several ARIMA imputations. Let's look at test sample #9 in MD1 for lambda = .15.

## Check the outliers in the ARIMA imputation.
```{r}
results = vector('list', 1)
idx = 1
for (j in list(10)) {
  # test data
  data_with_na = missing_list_md1[[2]][[j]]$data
  data_true = md1_df[missing_list_md1[[2]][[j]]$row:(missing_list_md1[[2]][[j]]$row+missing_list_md1[[2]][[j]]$size-1), ]$PWM_30min_avg
  na_true = data_true[missing_list_md1[[2]][[j]]$na.ind]
  
  # Impute the missing values using ARIMA model and Kalman smoothing
  imp_arima <- na.kalman(data_with_na, model = "auto.arima")

  # Extract all the imputed na values
  na_imp_arima = imp_arima[missing_list_md1[[2]][[j]]$na.ind]

  # compute rmse
  rmse_arima = sqrt(mean((na_imp_arima - na_true)^2))

  #Update results
  results[[idx]] <- c(lambda=missing_list_md1[[2]]$lambda, rmse=rmse_arima, imp=list(imp_arima), data.withNA=list(data_with_na))
  idx = idx + 1
}  
```


```{r plot_arima_imp_outliers1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = results[[1]]$data.withNA,
                   x.withImputations = results[[1]]$imp,
                   main = "Imputed Values with High RMSE Using ARIMA Model with Kalman Smoothing",
                   ylim = c(-10, 500), cex = .1)
```

Look into the variation in RMSE for structural model. E.g. for lambda=.1, sample 9 RMSE more than twice that of sample 1.
```{r}
results = vector('list', 2)
idx = 1
for (j in list(2, 9)) {
  # test data
  data_with_na = missing_list_md1[[2]][[j]]$data
  data_true = md1_df[missing_list_md1[[2]][[j]]$row:(missing_list_md1[[2]][[j]]$row+missing_list_md1[[2]][[j]]$size-1), ]$PWM_30min_avg
  na_true = data_true[missing_list_md1[[2]][[j]]$na.ind]
  
  # Impute the missing values using structural model and Kalman smoothing
  imp_struc <- na.kalman(data_with_na)

  # Extract all the imputed na values
  na_imp_struc = imp_struc[missing_list_md1[[2]][[j]]$na.ind]

  # compute rmse
  rmse_struc = sqrt(mean((na_imp_struc - na_true)^2))

  #Update results
  results[[idx]] <- c(lambda=missing_list_md1[[2]]$lambda, rmse=rmse_struc, imp=list(imp_struc), data.withNA=list(data_with_na))
  idx = idx + 1
}  
```


```{r plot_struc_imp_outliers1, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = results[[1]]$data.withNA[1:500],
                   x.withImputations = results[[1]]$imp[1:500],
                   main = "Imputed Values with low RMSE Using Structural Model with Kalman Smoothing",
                   ylim = c(250, 500), cex = .1)
```


```{r plot_struc_imp_outliers2, fig.height = 5, fig.width = 10}
plotNA.imputations(x.withNA = results[[2]]$data.withNA[1:500],
                   x.withImputations = results[[2]]$imp[1:500],
                   main = "Imputed Values with higher RMSE Using Structural Model with Kalman Smoothing",
                   ylim = c(250, 500), cex = .1)
```

In the above plots, sample 9 shows more variation in the data than in sample 1. This may account for the higher RMSE.

## Check for significance in the difference in the means of the RMSE.

### One-way ANOVA

#### Determine whether the means from the 4 imputation methods are equal.
```{r}
results_long_df <- reshape(results_df, varying = c('struc', 'arima', 'spline', 'ma'),
                           timevar = "method", v.names='RMSE',
                           times=c('struc', 'arima', 'spline', 'ma'), direction="long")
results_long_df$method <- as.factor(results_long_df$method)
results_long_df
```


```{r boxplots}
for (i in list(.1, .15, .2, .25)) {
  plot(RMSE ~ method, data = results_long_df[results_long_df$lambda == i, ],
       main = paste("Combined Buildings RMSE for Lambda = ", i))
}
```

#### Arima has outliers for all lambda values. Leave it out in the analysis.
```{r boxplots_wo_arima}
results_long_no_arima_df = results_long_df[results_long_df$method != "arima",]
results_long_no_arima_df$method <- factor(results_long_no_arima_df$method)
for (i in list(.1, .15, .2, .25)) {
plot(RMSE ~ method, data = results_long_no_arima_df[results_long_no_arima_df$lambda == i, ],
     main = paste("Combined Buildings RMSE for Lambda = ", i))
}
```

```{r}
for (i in list(.1, .15, .2, .25)) {
  cat("lambda = ", i, "\n")
  res = aov(RMSE ~ method, data = results_long_no_arima_df[results_long_no_arima_df$lambda == i, ])
  print(summary(res))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```

The anova tests found **insufficient evidence** to show that the differences are significant for all lambda values.

### Pairwise t-test for lambda .1, .15, .2, .25 (exclude ARIMA)

Pairwise multiple comparisons tests involve the computation of a p-value for each pair of the compared group. The p-value represents the risk of stating that an effect is statistically significant while this is not true. As the number of pairwise comparisons increases, and therefore the number of p-values, it becomes more likely to detect significant effects which are due to chance in reality.

To deal with this problem, multiple pairwise comparisons tests involve p-value corrections: p-values are penalized (= their value is increased) as the number of comparisons increase.

#### Results Summary
No significant pairwise differences found.

```{r}
for (i in list(.1, .15, .2, .25)) {
  cat("lambda = ", i, "\n")
  print(pairwise.t.test(results_long_no_arima_df[results_long_no_arima_df$lambda == i,]$RMSE,
                        results_long_no_arima_df[results_long_no_arima_df$lambda == i,]$method,
                        p.adjust = "bonferroni"))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```

for lambda = .15, exclude arima from test
```{r}
print(pairwise.t.test(results_long_df[(results_long_df$lambda == .15) & (results_long_df$method != "arima"),]$RMSE,
                      results_long_df[(results_long_df$lambda == .15) & (results_long_df$method != "arima"),]$method,
                      p.adjust = "bonferroni"))
```

## MD1 - Check for significance in the difference in the means of the RMSE.

### One-way ANOVA

#### Determine whether the means from the 4 imputation methods are equal.

```{r}
results_md1_df = as.data.frame(matrix(unlist(results_md1), nrow = length(results_md1), byrow = TRUE))
colnames(results_md1_df) = c("lambda", "struc", "arima", "spline", "ma")
results_long_md1_df <- reshape(results_md1_df, varying = c('struc', 'arima', 'spline', 'ma'),
                           timevar = "method", v.names='RMSE',
                           times=c('struc', 'arima', 'spline', 'ma'), direction="long")
results_long_md1_df$method <- as.factor(results_long_md1_df$method)
results_long_md1_df
```


```{r boxplots_md1}
for (i in list(.1, .15, .2, .25)) {
  plot(RMSE ~ method, data = results_long_md1_df[results_long_md1_df$lambda == i, ],
       main = paste("MD1 RMSE for Lambda = ", i))
}
```


```{r boxplots_wo_arima_md1}
results_long_no_arima_md1_df = results_long_md1_df[results_long_md1_df$method != "arima",]
results_long_no_arima_md1_df$method <- factor(results_long_no_arima_md1_df$method)
for (i in list(.1, .15, .2, .25)) {
plot(RMSE ~ method, data = results_long_no_arima_md1_df[results_long_no_arima_md1_df$lambda == i, ],
     main = paste("MD1 RMSE for Lambda =", i))
}
```

```{r}
for (i in list(.1, .15, .2, .25)) {
  cat("lambda = ", i, "\n")
  res = aov(RMSE ~ method, data = results_long_no_arima_md1_df[results_long_no_arima_md1_df$lambda == i, ])
  print(summary(res))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```

The anova tests found **insufficient evidence** to show that the differences are significant for all lambda values.

### Pairwise t-test for lambda .1, .15, .2, .25 (exclude ARIMA)

#### Results Summary
No significant pairwise differences found.
```{r}
for (i in list(.1, .15, .2, .25)) {
  cat("lambda = ", i, "\n")
  print(pairwise.t.test(results_long_no_arima_md1_df[results_long_no_arima_md1_df$lambda == i,]$RMSE,
                        results_long_no_arima_md1_df[results_long_no_arima_md1_df$lambda == i,]$method,
                        p.adjust = "bonferroni"))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```


## MD6 - Check for significance in the difference in the means of the RMSE.

### One-way ANOVA

#### Determine whether the means from the 4 imputation methods are equal.

```{r}
results_md6_df = as.data.frame(matrix(unlist(results_md6), nrow = length(results_md6), byrow = TRUE))
colnames(results_md6_df) = c("lambda", "struc", "arima", "spline", "ma")
results_long_md6_df <- reshape(results_md6_df, varying = c('struc', 'arima', 'spline', 'ma'),
                           timevar = "method", v.names='RMSE',
                           times=c('struc', 'arima', 'spline', 'ma'), direction="long")
results_long_md6_df$method <- as.factor(results_long_md6_df$method)
results_long_md6_df
```


```{r boxplots_md6}
for (i in list(.1, .15, .2, .25)) {
  plot(RMSE ~ method, data = results_long_md6_df[results_long_md6_df$lambda == i, ],
       main = paste("MD6 RMSE for Lambda = ", i))
}
```

```{r boxplots_wo_arima_md6}
results_long_no_arima_md6_df = results_long_md6_df[results_long_md6_df$method != "arima",]
results_long_no_arima_md6_df$method <- factor(results_long_no_arima_md6_df$method)
for (i in list(.1, .15, .2, .25)) {
plot(RMSE ~ method, data = results_long_no_arima_md6_df[results_long_no_arima_md6_df$lambda == i, ],
     main = paste("MD6 RMSE for Lambda = ", i))
}
```

```{r}
for (i in list(.1, .15, .2, .25)) {
  cat("lambda = ", i, "\n")
  res = aov(RMSE ~ method, data = results_long_no_arima_md6_df[results_long_no_arima_md6_df$lambda == i, ])
  print(summary(res))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```

The anova tests found **evidence** to show that the differences are significant for all lambda values.

### Pairwise t-test for lambda .1, .15, .2, .25 (exclude ARIMA)

#### Results Summary
for lambda=.1, ma > struc
for lambda=.15, ma > struc, spline
for lambda=.2, ma, spline > struc
for lambda=.25, ma > struc, spline

```{r}
for (i in list(.1, .15, .2, .25)) {
  cat("lambda = ", i, "\n")
  print(pairwise.t.test(results_long_no_arima_md6_df[results_long_no_arima_md6_df$lambda == i,]$RMSE,
                        results_long_no_arima_md6_df[results_long_no_arima_md6_df$lambda == i,]$method,
                        p.adjust = "bonferroni"))
  cat("--------------------------------------------------------------------------------------------------------------\n")
}
```

#### Show all the boxplots together 
```{r boxplots_grid}
par(mfrow = c(3, 4), mar = c(2, 2, 2, 2))
for (i in list(.1, .15, .2, .25)) {
plot(RMSE ~ method, data = results_long_no_arima_df[results_long_no_arima_df$lambda == i, ],
     main = paste("All Bldgs RMSE for λ = ", i), ylim = c(2, 22),
     xlab = NULL, ylab = NULL, cex.main = .7, cex.axis = .7)
}
for (i in list(.1, .15, .2, .25)) {
plot(RMSE ~ method, data = results_long_no_arima_md1_df[results_long_no_arima_md1_df$lambda == i, ],
     main = paste("MD1 RMSE for λ =", i), ylim = c(2, 22),
     xlab = NULL, ylab = NULL, cex.main = .7, cex.axis = .7)
}
for (i in list(.1, .15, .2, .25)) {
plot(RMSE ~ method, data = results_long_no_arima_md6_df[results_long_no_arima_md6_df$lambda == i, ],
     main = paste("MD6 RMSE for λ = ", i), ylim = c(2, 22),
     xlab = NULL, ylab = NULL, cex.main = .7, cex.axis = .7)
}
```

The distributions of RMSE by method in MD1 boxplot has less overlap which explains why some of the differences between methods are significant.


