---
title: "KE5105 - Building Electrical Consumption Forecasting"
output: github_document
---

# Extract, Transform and Load Data 7 - Data Imputation
### Gap Size Analysis for MD1 and MD6

# Summary of Findings
* RMSE increases with gap size.

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

# Load libraries
```{r}
library(ggplot2)
library(reshape2)
library(xts)
library(imputeTS)

source("/home/tkokkeng/Documents/KE5105/ETL/source/R/ETL.utils.R")
```

# Load data
```{r load_data, results='hide'}
# time series data
md1_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/processed_bldg_data/MD1.csv", header = TRUE, stringsAsFactors = FALSE)
md6_df <- read.csv("/home/tkokkeng/Documents/KE5105/ETL/source/processed_bldg_data/MD6.csv", header = TRUE, stringsAsFactors = FALSE)

# A list of 10 time series datasets with simulated missing data
missing_list_md1 <- readRDS("/home/tkokkeng/Documents/KE5105/ETL/source/md1-data/sim_missing_data_gapsize.md1.rds")
missing_list_md6 <- readRDS("/home/tkokkeng/Documents/KE5105/ETL/source/md6-data/sim_missing_data_gapsize.md6.rds")
```


```{r}
cat("length MD1 data = ", length(missing_list_md1), "\n", "length MD6 data = ", length(missing_list_md6))
```


```{r}
# Each list has lambda, missing1, ..., missing<n>
#nresults_md1 <- length(missing_list_md1) * (length(missing_list_md1[[1]]) - 2)
#nresults_md6 <- length(missing_list_md6) * (length(missing_list_md6[[1]]) - 2)
#results_md1 <- computeResultsWithGapsize("md1", nresults_md1, missing_list_md1, md1_df)
#results_md6 <- computeResultsWithGapsize("md6", nresults_md6, missing_list_md6, md6_df)
results_md1 <- computeResults("md1", missing_list_md1, md1_df, FALSE)
results_md6 <- computeResults("md6", missing_list_md6, md6_df, FALSE)
results <- c(results_md1, results_md6)
```


```{r}
results_df = as.data.frame(matrix(unlist(results), nrow = length(results), byrow = TRUE))
colnames(results_df) = c("lambda", "gapsize", "na_percent", "struc", "arima", "spline", "ma")
results_df
```

Save results to file. 
```{r}
write.csv(results_df, file = "results_gapsize.csv", row.names = FALSE)
```


```{r}
agg_df <- aggregate(results_df[, c("na_percent", "struc", "arima", "spline", "ma")],
                    list(lambda = results_df$lambda, gapsize = results_df$gapsize), mean)
agg_df
```


```{r}
agg_long_df <- melt(agg_df, id.vars = c("lambda", "gapsize", "na_percent"), variable.name = "method")
colnames(agg_long_df)[5] <- "RMSE"
agg_long_df
```

RMSE appears to increase with gap size and / or the percentage of NAs.
```{r rmse_gapsize_plot, fig.height = 10, fig.width = 7}
ggplot(agg_long_df, aes(x=gapsize, y=RMSE, group=method)) +
  geom_line(aes(color=method)) +
  geom_point(aes(color=method, size=na_percent), shape = 1) +
  #scale_size_continuous(range = c(.2,10), trans = "exp") +
  facet_grid(lambda ~ .) +
  ggtitle("Imputation RMSE by Gap Size and Lambda")
```


```{r rmse_na_percent_plot, fig.height = 10, fig.width = 7}
ggplot(agg_long_df, aes(x=na_percent, y=RMSE, group=method)) +
  geom_line(aes(color=method)) +
  geom_point(aes(color=method, size=gapsize), shape = 1) +
  scale_size_continuous(range = c(1,10)) +
  facet_grid(lambda ~ .) +
  ggtitle("Imputation RMSE by NA percentage and Lambda")
```

For a constant gap size, the error does not increase with the total missing data (or the frequency of missing data).

```{r rmse_na_percent_gapsize_plot, fig.height = 10, fig.width = 7}
ggplot(agg_long_df, aes(x=na_percent, y=RMSE, group=method)) +
  geom_line(aes(color=method)) +
  geom_point(aes(color=method), shape = 1, size = 1) +
  # geom_point(aes(color=method, size=gapsize), shape = 1) +
  # scale_size_continuous(range = c(1,10)) +
  facet_grid(gapsize ~ .) +
  ggtitle("Imputation RMSE by NA percentage and Gap Size")
```

The percentage of NAs increases linearly for small gap sizes and exponentially for larger gap sizes.

```{r na_percent_gapsize_plot, fig.height = 10, fig.width = 7}
ggplot(agg_long_df, aes(x=na_percent, y=lambda)) +
  geom_line() +
  geom_point(shape = 1, size = 1) +
  # geom_point(aes(color=method, size=gapsize), shape = 1) +
  # scale_size_continuous(range = c(1,10)) +
  facet_grid(gapsize ~ .) +
  ggtitle("NA percentage by Gap Size")
```
